---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

## Overview

The **bmrr** package fits a Bayesian Multivariate Rank Regression (BMRR). The BMRR model serves to aggregate multivariate ranking data in cases where raters rank multiple items over multiple time points or criteria. Using latent hierarchical regression models, BMRR pools information across raters and criteria, incorporates covariates and rater expertise, and constructs an aggregate ranking. Crucially, BMRR provides full posterior uncertainty quantification for this aggregate ranking and all model parameters, and readily handles partial (i.e., incomplete) rankings and ties. In addition to fitting the model, the package provides helpful posterior distribution plots to analyze aspects of the aggregated ranking, the raters themselves, the covariate effects, and provides comparisons across items.

## How to use bmrr

```{r}
#--- Load in package
library(bmrr)

#--- Load in some additional packages
library(dplyr)
library(ggplot2)
library(tidyr)
```

Below we walk through an applied example with Major League Baseball (MLB) ranking data.

```{r}
#--- Load in the data
data(sports_data)

# Define Simulation Starting values & Priors ----------------------------------

# Sampling Specifications ----------------------------------------------------
#--- Set seed
seed = 24
set.seed(seed)

#--- Number of samples
samps = 20000

#--- Burn-in
burnin = 4000

#--- number of items 
num_items = m = length(unique(sports_data$player_id))

#--- number of raters
num_raters = n = 5

#--- number of times
num_criteria = t = length(unique(sports_data$Year))

#--- number of covariates
c = 3

#--- Priors for beta and lambda
lambda = beta = .8

#--- Yes Covars are Present
covars_present = TRUE


# Clean Data & Parameters --------------------------------------------

covars = array(NA,dim = c(m,c,t))

for(i in 1:t){
  covars[,,i] <- sports_data %>%
    arrange(player_id)%>%
    filter(Year == 2020+i)%>%
    ungroup()%>%
    dplyr::select(
      player_age, 
      WAR, 
      base_salary,
    )%>%as.matrix()
}

sports_data_array <- array(NA,dim = c(n,m,t))
for(i in 1:t){
  sports_data_array[,,i] <- sports_data%>%
    arrange(player_id)%>%
    filter(Year == 2020+i)%>%
    ungroup()%>%
    dplyr::select(
      ESPN, 
      CBS,
      MLB, 
      BR, 
      Yahoo
    )%>%t()
}

data <- sports_data_array

```

```{r}
# Priors -----------------------------------------------------------------------

#--- Prior for xi (hierarchical mean)
mu_0 = rep(0, m) #0
sigma_0 = rep(1,m) #1

#--- Prior for alpha (hierarchical variance)
alpha_0 <- .8

#--- Prior for beta
beta_0 = rep(0, c+1) #there are 'c' coefficients 
sum_0 = diag(100,c+1) #there are 'c' coefficients
sum_0_inverse = solve(sum_0)

#--- Prior for Sigma
lambda = .8


# Run Simulation ------------------------------------------------------------
posterior_output <- run_bmrr(num_raters,
                     num_items,
                     num_criteria,
                     data,
                     samps,
                     burnin,
                     covars_present,
                     covars,
                     
                     mu_0,
                     sigma_0,
                     alpha_0,
                     beta_0,
                     sum_0_inverse,
                     lambda,
                     seed
)


# Simulation Results------------------------------------------------------------

#--- Mu Posterior
mu_posterior <- posterior_output$mu_posterior

#--- Xi Posterior
xi_posterior <- posterior_output$xi_posterior

#--- Sigma Posterior
sigma_posterior <- posterior_output$sigma_posterior

#--- Alpha Posterior
alpha_posterior <- posterior_output$alpha_posterior 

#--- Beta Posterior
beta_posterior <- posterior_output$beta_posterior

#--- Z Posterior
Z_posterior <- posterior_output$Z_posterior

#--- Run Time
time <- as.numeric(posterior_output$time)

```

```{r, eval = FALSE}

mu_beta_plot_list <- mu_beta_plots(mu_posterior, 
                         beta_posterior,
                         covars,
                         item_names, 
                         criteria_names)

mu_beta_plot_list[[5]]
mu_beta_plot_list[[6]] # MAIN PLOT FOR PAPER

```

```{r, eval }

## Posterior Xi ----------------------------------------------------------------
xi_plot_list <- xi_plots(xi_posterior, 
                         item_names)

xi_plot_list[[1]] 
xi_plot_list[[2]] # PLOT FOR PAPER

```

```{r}

covariate_names <- c( "Age", "WAR", "Salary")


beta_plots_list <- beta_plots(beta_posterior, 
                              covariate_names)

beta_plots_list[[2]] # plot for paper


```

```{r}
ranker_names <-c("ESPN","CBS", "MLB","BR","Yahoo")
colMeans(sigma_posterior)
colSds(sigma_posterior)

sigma_plots_list <- sigma_plots(sigma_posterior, 
                                ranker_names)
sigma_plots_list

```

```{r}

time1 = (mu_posterior[, ,4] + (beta_posterior%*%t(covars[,,4]))) %>% 
  as.data.frame()%>%
  mutate(
    iter = row_number()
  )

compare_prob_pos_table = matrix(NA, nrow = m, ncol = m)
#--- Reading Column to Row (Probability Column 1 is larger than Row 2)

for(j in 1:m){
  for(k in 1:m){
    prob_now = sum(time1[,j] >= time1[,k])/dim(time1)[1]
    if(j == k){
      compare_prob_pos_table[k,j] = 1
    }
    else{
      compare_prob_pos_table[k,j]<- prob_now
    }
  }
}


compare_prob_pos_table <- data.frame(compare_prob_pos_table)%>%
  mutate_if(is.character,as.numeric)

names(compare_prob_pos_table) <-item_names

rownames(compare_prob_pos_table) <- item_names

compare_prob_pos_table = round(compare_prob_pos_table, 3)

compare_prob_pos_table_OF = compare_prob_pos_table %>% 
  dplyr::select(

    #--- Outfield
    "George Springer",
    "Mike Trout",          
    "Bryce Harper",
    "Yordan Alvarez",
    "Aaron Judge",
    "Mookie Betts",
    "Kyle Tucker",
    "Christian Yelich",
    "Teoscar Hernandez" ,
    "Brandon Nimmo",
    "Ronald Acuna Jr",
    "Fernando Tatis Jr",
    "Juan Soto" ,
    "Randy Arozarena" ,
    "Luis Robert Jr" 

  )%>%
  filter(
    row.names(compare_prob_pos_table) %in% c("George Springer",
                       "Mike Trout",          
                       "Bryce Harper",
                       "Yordan Alvarez",
                       "Aaron Judge",
                       "Mookie Betts",
                       "Kyle Tucker",
                       "Christian Yelich",
                       "Teoscar Hernandez" ,
                       "Brandon Nimmo",
                       "Ronald Acuna Jr",
                       "Fernando Tatis Jr",
                       "Juan Soto" ,
                       "Randy Arozarena" ,
                       "Luis Robert Jr" )
  )


compare_prob_pos_table_OF <- (as.matrix(compare_prob_pos_table_OF))


sorted_row_names <- sort(rownames(compare_prob_pos_table_OF))
sorted_col_names <- rev(sort(colnames(compare_prob_pos_table_OF)))


long_data <- melt(compare_prob_pos_table_OF, varnames = c("RowItem", "ColumnItem"), value.name = "Probability")

long_data$RowItem <- factor(long_data$RowItem, levels = sorted_row_names)
long_data$ColumnItem <- factor(long_data$ColumnItem, levels = sorted_col_names)

#--- Create the Heatmap

long_data = long_data %>% 
  mutate(Probability = 1- Probability)%>%
  mutate(Probability = ifelse(RowItem == ColumnItem, 1, Probability))

# Create the base plot
heatmap_plot <- ggplot(
  data = long_data,
  aes(x = ColumnItem, y = RowItem, fill = Probability)
) +
  geom_tile(color = "white")+  
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 10),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, face = "bold"), 
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    axis.ticks = element_blank()
  )+ 
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = .5,
    limit = c(0, 1)
  ) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank()
  )

heatmap_plot


```

## Installation

```{r, eval = FALSE}
if (!requireNamespace("remotes")) {
  install.packages("remotes")
}
remotes::install_github("Rose-Graves/bmrr")
```
