---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

## Overview

The **bmrr** package fits a Bayesian Multivariate Rank Regression (BMRR). The BMRR model serves to aggregate multivariate ranking data in cases where raters rank multiple items over multiple time points or criteria. Using latent hierarchical regression models, BMRR pools information across raters and criteria, incorporates covariates and rater expertise, and constructs an aggregate ranking. Crucially, BMRR provides full posterior uncertainty quantification for this aggregate ranking and all model parameters, and readily handles partial (i.e., incomplete) rankings and ties. In addition to fitting the model, the package provides helpful posterior distribution plots to analyze aspects of the aggregated ranking, the raters themselves, the covariate effects, and provides comparisons across items.

## How to use bmrr

```{r, eval = FALSE}
library(bmrr)
```

Below we walk through a simple simulated example using most of the functions within bmrr. Within this example we have 10 items and 8 raters ranking them over 3 time points with some partial rankings (some raters not ranking all the items at every time point).

For an applied example on sports ranking data please see the vignettes folder.

```{r, eval = FALSE}

  # Sampling Specifications ----------------------------------------------------
  #--- Set seed
  seed = 24
  set.seed(seed)
  
  #--- Number of samples
  samps = 4000
  
  #--- Burn-in samples
  burnin = 1000
  
  # Data Generation ------------------------------------------------------------
  #--- number of items 
  num_items = m =  10
  
  #--- number of raters
  num_raters = n = 8
  
  #--- number of times
  num_criteria = K = 3
  
  #--- number of covariates
  num_covars = c = 2
  covars_present = TRUE
  
  #--- variance between raters/items to be the same
  same_alpha = TRUE 
  same_sigma = TRUE
  
  #--- Missingness levels in the data
  partial_raters = 4
  missing_percent = 0.3
  
  #--- Simulate the data & parameters
  sim_data_parms <- simulate_data(num_items,     
                                  same_alpha,
                                  num_criteria,
                                  num_raters,    
                                  same_sigma,
                                  num_covars,
                                  partial_raters,
                                  missing_percent,
                                  seed
  )
  
  
  data <- sim_data_parms$data
  covars <- sim_data_parms$covars
```

```{r, eval = FALSE}
  # Priors -----------------------------------------------------------------------
  
  #--- Prior for xi (hierarchical mean)
  mu_0 = 0
  sigma_0 = 1
  
  #--- Prior for alpha (hierarchical variance)
  alpha_0 = .8
  
  #--- Prior for beta
  beta_0 = rep(0, c+1) #there are 'c' coefficients 
  sum_0 = diag(100,c+1) #there are 'c' coefficients
  sum_0_inverse = solve(sum_0)
  
  #--- Prior for Sigma
  lambda = .8
  
  # Running Simulation -----------------------------------------------------------
  posteior_output <- run_bmmr(num_raters,
                              num_items,
                              num_criteria,
                              data,
                              samps,
                              burnin,
                              covars_present,
                              covars,
                              
                              mu_0,
                              sigma_0,
                              alpha_0,
                              beta_0,
                              sum_0_inverse,
                              lambda,
                              
                              seed
  )
  
  # Output Options ---------------------------------------------------------------
  
  #--- Mu Posterior
  mu_posterior <- posteior_output$mu_posterior
  
  #--- Xi Posterior
  xi_posterior <- posteior_output$xi_posterior
  
  #--- Sigma Posterior
  sigma_posterior <- posteior_output$sigma_posterior
  
  #--- Alpha Posterior
  alpha_posterior <- posteior_output$alpha_posterior 
  
  #--- Beta Posterior
  beta_posterior <- posteior_output$beta_posterior
  
  #--- Z Posterior
  Z_posterior <- posteior_output$Z_posterior
  
  #--- Time as an integer
  time <- as.numeric(posteior_output$time)


```

## Installation

```{r, eval = FALSE}
if (!requireNamespace("remotes")) {
  install.packages("remotes")
}
remotes::install_github("Rose-Graves/bmrr")
```
